Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-02-10 article/article.Rmd
+ include random prioritisations to verify surrogate performance + change dependnecies MAKEFILE to avoid re-running stuff + rewrite of results in article + changes to stats

## Statistical analysis
We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models [NLMMs; using the $\texttt{lme4}$ R package; @r469]. A model was fit to the genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = a - (a - 0) \times exp(-r \times x)$; where $a$ is the asymptote and $r$ is the accumulate rate). An interaction term was included to estimate parameters separately for adaptive and neutral variation. The models accommodated species-level variation using random effects for both parameters.

We tested for differences in the effectiveness of prioritizations generated under different scenarios and targets using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package]. We fit a full model to the proportion of genomic variation secured in all prioritizations using a logit link. Both models contained variables indicating the scenario used to generate the prioritization, the targets used to generate the prioritizations, and the type of genetic variation measured. The model was subject to a step-wise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$). All models contained an intercept-only random effect to ensure that comparisons were made within species. Finally, we conducted a post-hoc analysis using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

## Results
<!-- Single species prioritizations -->
Surrogates-targets improved the performance of single-species prioritizations (Figure 1). There were significant differences among the performance of prioritizations depending on the targets used to generate them and if they were assessed according to adaptive or neutral genetic variation (${\chi^{2}}_{`r single.spp.AOV1[['Df']][1]`,`r single.spp.AOV1[['Df']][2]`} = `r round(single.spp.AOV1[['Chisq']][2],2)`$, $P `r pretty.pval(single.spp.AOV1[['Pr(>Chisq)']][2])`$). Specifically, the amount-based prioritization captured much less adaptive and neutral genetic variation than either the surrogate-based ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Neutral variation - Amount.Neutral variation']])`$) or genetic-based ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Amount.Neutral variation']])`$) prioritizations. No differences were detected in the proportion of adaptive ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Surrogate.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Surrogate.Adaptive variation']])`$) or neutral genetic variation ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Surrogate.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Surrogate.Neutral variation']])`$) secured by the the surrogate-based and genetic-based prioritizations. However, these patterns were not reflected in the multi-species prioritizations.

<!-- Multi-species prioritizations -->
Environmental and geographic surrogates were not found to improve the performance of multi-species prioritizations (Figure 2; Figure 3). No significant differences in the effectiveness of prioritizations were found depending on whether they were generated using different targets (${\chi^{2}}_{`r multi.spp.AOV3[['Df']][1]`,`r multi.spp.AOV3[['Df']][2]`} = `r round(multi.spp.AOV3[['Chisq']][2],2)`$; $P `r pretty.pval(multi.spp.AOV3[['Pr(>Chisq)']][2])`$), if they were assessed using adaptive or neutral genetic variation (${\chi^{2}}_{`r multi.spp.AOV2[['Df']][1]`,`r multi.spp.AOV2[['Df']][2]`} = `r round(multi.spp.AOV2[['Chisq']][2],2)`$; $P `r pretty.pval(multi.spp.AOV2[['Pr(>Chisq)']][2])`$), nor the interaction between these variables (${\chi^{2}}_{`r multi.spp.AOV1[['Df']][1]`,`r multi.spp.AOV1[['Df']][2]`} = `r round(multi.spp.AOV1[['Chisq']][2],2)`$; $P  `r pretty.pval(multi.spp.AOV1[['Pr(>Chisq)']][2])`$).

<!-- Correlation analysis -->
The surrogate variables showed a greater utility for neutral genetic variation than adaptive variation (Figure 4). The accumulation rate ($r$) was `r round(summary(env.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(env.correlation.NLMM)$coefficients[2],2)` S.E. for the proportion of adaptive variation captured in a prioritization using environmental surrogates. On the other hand, the rate ($r$) was `r round(summary(geo.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(geo.correlation.NLMM)$coefficients[2],2)` S.E. for neutral variation using geographic surrogates. Thus minor increases in the target for either surrogate results in major increases in the benefit of the prioritization. However, the coefficient was much larger for neutral genetic variation so minor increases in the target can achieve substantial gains.

>>>>

## Statistical analysis
We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469]. All models were fit to the proportion of genetic variation that prioritizations secured using a logit link. To understand the relationship between the surrogate-based target used to generate a prioritization and how much genetic variation the resulting prioritization secured, we fit a full model containing a continuous quadratic predictor variable for the surrogate target used to generate the prioritization, a categorical variable indicating the type of genetic variation, and a variable for the interaction between the previous variables. This model contained random effects for both the intercept and the predictor variables to accommodate variation between species. To determine if surrogate-based targets improved the effectiveness of prioritizations under different scenarios, we fit a full model containing categorical variables indicating the scenario used to generate the prioritization, the combination of targets used to generate the prioritizations, and the type of genetic variation measured. This model contained intercept-only random effects. Both models were subject to step-wise term deletion routines using likelihood ratio tests (LRT) to assess significance (terms were retained when $P < 0.05$). Finally, we conducted a post-hoc analysis using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

## Results
<!-- Surrogacy analysis -->
Prioritizations generated using greater environmental and geographic surrogate-based targets were found to secure greater proportions of adaptive and neutral genetic variation (Figure 1). The proportion of genetic variation that these prioritizations secured was substantially more than the genetic variation secured by randomly generated prioritizations (`r xxx`). Generally, prioritizations needed to capture at least `r xxx` % of the environmental and geographic surrogates to secure 90 % of the species' genetic variation. Additionally, we detected differences in the effectiveness of the environmental and geographic surrogates (`r xxxx`). Increases in environmental surrogate-based targets resulted prioritizations that secured more adaptive genetic variation, than geographic surrogate-based targets for neutral variation. These results demonstrate that environmental and geographic variables are effective surrogates for genetic variation.

<!-- Scenario analysis -->
The addition of environmental and geographic surrogate-based targets to amount-based targets was not found to substantially improve the effectiveness of prioritizations under any of the scenarios (Figure 2). No significant differences were detected between prioritizations generated using different combinations of targets (${\chi^{2}}_{`r `,`r `} = `r `$, $P `r `$), or under any of the different scenarios (${\chi^{2}}_{`r `,`r `} = `r `$, $P `r `$). The only difference was that the inclusion of surrogate-based targets tended to result in prioritizations that secured slightly more adaptive than neutral genetic variation. Under the simplest of circumstances, the single-species amount-based prioritizations secured most of the species' adaptive (mean `r ` \pm `r `) and neutral (mean `r ` \pm `r `) genetic variation. These proportions were not significantly different to those associated with prioritizations generated using additional surrogate-based targets ($Z = `r `$, $P `r `$) or genetic-based targets ($Z = `r `$, $P `r `$). We found similar patterns in the multi-species scenario () and the multi-species with opportunity cost scenario (). Although the prioritizations had similar levels of performance, there were different in the their spatial distributions (Figure 3). In particular, the inclusion of cost data in the multi-species prioritizations had the effect of concentrating conservation effort towards areas with higher elevation. Overall, these results suggest that while environmental and geographic surrogate variables are effective surrogates, they are not needed to generate prioritizations that secure most of the intra-specific genetic variation.
